---
title: "R basics: Data types and programming skills"
author: "Theresa Gessler and Hauke Licht"
date: "last updated: `r format(Sys.Date(), '%e %B %Y')`"
description: >
  Refresh your basic R programming skills.
  This tutorial reviews
  common R data types,
  how to create, access and modify objects,
  how to loop/iterate,
  and 
  how to write functions.
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---


```{r knitr, include=FALSE}
knitr::opts_chunk$set(
  # code chunk options
  echo = TRUE
  , eval = TRUE
  , warning = FALSE
  , message = FALSE
  , cached = FALSE 
  , exercise = TRUE
  , exercise.completion	= TRUE
)
```

```{r setup-invisble, exercise = FALSE, include=FALSE}
library(learnr)
library(dplyr)
# library(rvest)
```


## Introduction


In this tutorial, we will refresh your basic R programming skills.
It covers some of the essential programming concepts we will need to scrape and clean web data. 
Specifically, we will review

- common R data types,
- how to create, access and modify objects,
- how to loop/iterate, and 
- how to write functions.

If you're having trouble to understand the code in other tutorials, we recommend you go through this tutorial and make sure you understand each of the concepts introduced.

## Creating objects (assignment)

To create an object in R, we use the `<-` operator.
For example, we can create an object called 'x' with the value 1 as follows:

```{r assignment}
x <- 1
```

This is called **assignment**.
So the expression `x <- 1` reads "I assign the value 1 to an object called 'x'."

<!-- Note that there is a keyboard shortcut for inserting this operator: alt + "-" (dash) -->

### Use `<-`, not `=` for assignment

Actually, you could also use `=` to assign values to objects.
However, it is an R code convention to reserve `=` for passing values to function arguments.
This makes reading code easier. 
For example:

```{r, exercise = FALSE}
res <- length(x = 1:3)
```

### Evaluating objects 

After assigning an object, it can be used by referring to its name.
(The technical term is "evaluation".) 

```{r evaluation}
# create 'x'
x <- 1
# "call"/"print" 'x' (i.e. evaluate it)
x
```

## Data types

R has many **data types** (see `?typeof`).
The most common (atomic) types are:

1. "logical",
2. "integer",
3. "double",
4. "character",
5. `NULL`

### 1. logical

The simplest data type is "logical".
There are only three valid logical values:  `TRUE`, `FALSE`, or `NA` ("missing").
You can check whether a value has type logical using `is.logical()`.

```{r type-logical}
typeof(TRUE)
typeof(FALSE)
typeof(NA)

is.logical(TRUE)
is.logical(FALSE)
is.logical(NA)
```

### 2. integer

Integer values are numbers: ..., -1, 0, 1, ...
You can check whether a value has type logical using `is.integer()`.
A missing integer can be created with `NA_integer_`.

Note that you have to explicitly force a number to type integer for R to recognize it as such.
This can be done by adding an `L` behind the number (e.g. `1L`).
You need to force this behavior, because R interprets numbers as "double" type by default.

```{r type-integer}
is.integer(1) # FALSE !!!
is.integer(1L) # TRUE =)
is.integer(NA_integer_) 
```


### 3. double

Integer values are *real* numbers in (-&infin;, &infin;)
You can check whether a value has type logical using `is.double()`.
A missing integer can be created with `NA_real_` (not `NA_double_`!!!).

```{r type-double}
typeof(1)
is.double(1)
is.double(NA_real_)
```

#### *Special values*

Note that because the double type implements real-valued mathematical operations in R, 
there are two special values reserved for *infinity* and *not-a-number* (undefined values):

```{r infinity}
is.double(Inf)
Inf
-Inf
```


```{r nan}
typeof(NaN)
is.na(NaN)

# the logical NA and NaN are two different pairs of shoes
identical(NA, NaN)
# the same applies to double-type NA
identical(NA_real_, NaN)
```

### 4. character

Character values are strings, for example "hello world!"
You can create a character value by wrapping it inside single or double quotation marks.
A missing character values can be created with `NA_character_`.
You can check whether a value has type character using `is.character()`.

```{r type-character}
is.character('this is a valid character values')
is.character("this is also a valid character values")
is.character(NA_character_)
```

### 5. `NULL`

`NULL` is an object that represents "no data."
Contrast this with the `NA*` values introduced above.

```{r null-type}
# in the case of NULL, there is really nothing!
length(NULL)

# NA values, in contrast, represent some data
# (the value is unknown, but this is considered valid information)
length(NA)
length(NA_integer_)
length(NA_real_)
length(NA_character_)
```

## Vectors

(Ordered) collections of objects of the same type are called **vectors**.
You probably know vectors from math in school.

```
1 # a "scalar"
(1, 2, 0) # a "vector"
```

In the above example, the vector has three **elements**: 1, 2, and 0

### Creating vectors

In R, you create vectors by **c**ombining logical, integer, double or character values using the `c()` function.

For example, you can combine the three valid logical values in a logical vector:

```{r creating-vectors}
# create a logical vector
logicals <- c(TRUE, FALSE, NA)
typeof(logicals)
is.logical(logicals)
```

### Type conversion 

Note however that you can always only combine objects of the same type in an (atomic) vector
If you violate this rule, R will convert all values to the type according to the following hierarchy: `logical` &#8614; `integer` &#8614; `double` &#8614;	`character`.

That is, if you combine a logical and an integer value, both will be represented as integers; 
if you combine a logical, an integer and a double value, all will be represented as double values;
and so on.

This is called **type conversion**.

```{r type-conversion}
# logical converts (upwards) to integer 
typeof(c(TRUE, 1L))
c(TRUE, 1L)

# logical and integer convert (upwards) to double 
typeof(c(TRUE, 1L, 1.1))
c(TRUE, 1L, 1.1)

# logical, integer and double convert (upwards) to character 
typeof(c(TRUE, 1L, 1.1, "one"))
c(TRUE, 1L, 1.1, "one")
```

### `NULL` is ignored 

Note that because `NULL` represents "no data," it is *ignored* when creating an atomic vector!

```{r null-in-vector, exercise.lines = 1}
c(NULL, "value")
```

Accordingly, it is also ignored when performing type conversion.

```{r null-in-type-conversion, exercise.lines = 1}
c(NULL, TRUE, 1L, 1.1, "one")
c("NULL", NULL)

```

### Naming vectors

One thing you haven't seen so far in this tutorial are **named vectors**.
A named vector is a vector where _each_ element has a corresponding character value that denotes its "name". 

You can create a named vector using name--value syntax, or with the `setNames()` function.
You can assigning names to an already existing vector using the `names()` function.

```{r naming-vectors}
# creating a named vector
c("apple" = 3, "banana" = 2, "lemon" = 1.5)
setName(c(3, 2, 1.5), c("apple", "banana", "lemon"))

# assigning names to an existing vector
capitals <- c("Lisbon", "Bejing", "Cape Town")
names(capitals) <- c("Portugal", "China", "South Afric")
```

You can access the names of a vector using the `names()` function.

```{r query-names, exercise.lines = 2}
prices <- c("apple" = 3, "banana" = 2, "lemon" = 1.5)
names(prices)
```

And because `names()` returns `NULL` if a vector is unnamed, you can it to check if a vector is named.

```{r check-if-named}
# create unnamed vector
x <- logical(3)
# is named?
is.null(names(x)) # `FALSE`

# name
names(x) <- c("a", "b", "c")
is.null(names(x)) # TRUE
```

### Extracting values from vectors

Once you have created a vector, you can access its values by **indexing** it.
To do so, use the `[` function and provide integer values to access positions as follows:

```{r integer-indexing}
fruits <- c("apple", "banana", "lemon")
# access first value
fruits[1]
# access first two values
fruits[1:2]
```

Note that by negating indexes, for example, by adding a negative sign ("dash") in front of an integer index value, you can **omit** ("drop") values:

```{r negative-indexing}
fruits <- c("apple", "banana", "lemon")
# omit first value
fruits[-1]
# omit first two values
fruits[-c(1:2)]
```

### Indexing with logical values

You can also access/omit the values of a vector using logical values.
(Note that `!` can be used to negate boolean values.)

```{r boolean-indexing}
idxs <- c(TRUE, FALSE, FALSE)
# access first value
fruits[idxs]
# negate to access last two values
fruits[!idxs]
```

### Indexing named vectors

Named vectors are useful because they allow indexing by names:
```{r name-indexing}
prices <- c("apple" = 3, "banana" = 2, "lemon" = 1.5)
prices["apple"]
prices[c("apple", "lemon")]

# NOTE: 
#  ... negative indexing doesn't work with names
prices[-c("apple", "lemon")] # this throws an ERROR!
#  ... but you can use logical indexing 
prices[!names(prices) %in% c("apple", "lemon")]
```


### *Empty/zero-length vectors* (optional)

Empty vectors can be created with the following functions:

```{r creating-empty-vectors}
logical()
integer()
double()
character()
```

Note that if you pass an positive numeric value to the `length` argument of these functions (the  default is `length = 0L`, see also `?vector`), 
you can create vectors of given length.

Note, however, that the initial values picked vary between types. 

```{r creating-fixed-length-vectors}
logical(3) # all values are `FALSE`
integer(3) # all values are `0L`
double(3) # all values are `0`
character(3) # all values are `""`
```

### *Other functions for creating vectors* (optional)

For certain types, there are shortcuts for creating vectors.
For example, `seq()` creates a sequence of numeric (double) values.
Specifically, `seq()` can be used to create sequence of numbers that in- or decreases by a specific amount specified by the `by` argument (the default is `by = 1L`).

```{r creating-increasing-sequences}
# increasing in steps of 1
seq(-10, 10) # equivalent to `-10:10` 

# increasing in steps of 5
seq(-10, 10, by = 5)

# NOTE: if you don't force the value passed to `by` to an integer, 
#  `seq()` will return a double vector
typeof(seq(-10, 10, by = 5))
typeof(seq(-10, 10, by = 5L))
```

You can also create decreasing sequences!
But remember to bring the start and end values in the correct order.

```{r creating-decreasing-sequences, exercise.lines = 1}
seq(-10, 10, by = -5L)
```

`seq()` can also be called with the argument `length.out`.
This allows creating a sequence of a pre-defined length.

```{r creating-fixed-length-sequences, exercise.lines = 1}
seq(-10, 10, length.out = 11)
```

Other useful functions are `seq_len()` and  `seq_along()`.
`seq_len()` takes a non-negative number as input and returns an integer sequence of equal length.

```{r seq_len, exercise.lines = 2}
seq_len(0)
seq_len(3)
```

`seq_along()` takes a vector as input and returns an integer sequence of equal length as the input vector.

```{r seq_along, exercise.lines = 2}
seq_along(c())
seq_along(c("a", "b", "c"))
```

## Atomic and recursive types

### Atomic types 

The data types you've encountered so far in this tutorial (logical, integer, double, character, and `NULL`) are **atomic types**.
They are "atomic" in the sense that they cannot "nest" objects of other type.

So if you use `c()` to create an atomic vector, all values need to have the *same* type.
Remember that if you violate this rule, R performs *type conversion*  following hierarchy: `logical` &#8614; `integer` &#8614; `double` &#8614;	`character`.

### Recursive types 

In contrast to atomic types, **recursive types** can nest objects of other types.

The most important recursive type is "list".

## Lists

You can create a list with the `list()` functions

```{r first-list}
list(TRUE, 1L, 1.1, "one")
```

### No type conversion inside lists

You'll note that *no* "type conversion" is performed inside lists!

```{r list-no-type-conversion}
# types are preserved
list(TRUE, 1L, 1.1, "one")
# types are  n o t  preserved
c(TRUE, 1L, 1.1, "one")
```

### Combining lists 

You can use `c()` to combine multiple list. 

```{r combining-lists, exercise.lines = 1}
c(list(1), list(2))
```

But note that when combining a list/lists and a vector/vectors, the vector(s) will be converted to a list(s)!
So **type conversion** applies beyond atomic vectors 

```{r combining-lists-type-conversion}
c(
  list("hello"), # a list with one character element
  c("goodbye") # a character vector (!) with one element
)
```

### Naming lists

Like vectors, lists can be named:

```{r naming-lists}
# create named list 
x <- list("logical" = TRUE, "integer" = 1L, "double" = 1.1, "character" = "one")
names(x)

# assign names to an existing list
x <- list(TRUE, 1L, 1.1, "one")
names(x) <- c("logical", "integer", "double", "character")
names(x)
```

### Indexing lists

The difference between atomic and recursive types is important because it matters for how you can **index** objects, that is, access their values.

As already shown, the elements of atomic vectors can be accessed with a single square bracket `[`.
In contrast, we use double square brackets `[[` to access the elements of a list.

```{r vector-vs-list}
vec <- c(1, 2)
vec[1]

lst <- list(1, 2)
lst[[1]]
```

This difference in syntax is necessary because, in contrast to atomic types, lists can be multiple levels deep.
This is illustrated by the example below:

```{r list-grandma}
# (a list of lists of lists)

# grandma Paula has two children: ...
paula <- list(
  # ... Frank
  "frank" = list(
    # Frank has two children ...
    "laura" = list(),
    "beth" = list()
  ),
  # ... and Tom
  "tom" = list(
    # ... Tom has no children
  )
)
```

To access elements at the first level (grandma Paula's children), we need to use `[[`.
For example, So to get the data for Paula' first child, Frank' we need to index the first element of the fist level.

```{r list-grandma-extract-l1}
# get data for Paula's first child
paula[[1]]

# alternatively:
paula[["frank"]]
paula$frank
```

And to get at the values at the second level, we need to index even deeper.
Below, we extract the information for the first child of Frank, who in turn is grandma Paula's first child.

```{r list-grandma-extract-l2}
# get data for first child of Paula's first child
paula[[1]][[1]]
```


## Data frames

In R, there are several **data structures** that combine multiple values into a single object.
You have already encountered two of the most important data structures in R:

- (atomic) vectors (`vector()`): sequence of values of a certain type
- lists (`list()`): collection of objects, potentially of different types

A third, very important data structure is the **data frame**. 
A data frame is a tabular data format similar to an Excel spread sheet or a CSV file.

### Creating data frames

Data frames allow us to combine many vectors or lists of the *same length* into a single object.
We create data frames with the `data.frame()` function.

```{r dataframes}
student_names <- c("Bill", "Jane", "Sarah", "Fred", "Paul")
math_scores <- c(80, 75, 91, 67, 56)
verbal_scores <- c(72, 90, 99, 60, 68)

students <- data.frame(student_names, math_scores, verbal_scores)
students
```

Note that `student_names` has a different type (character) than `math_scores` (numeric).
Nevertheless, you can combine them in a data frame. 
(This is because a data frame is really just a list of lists with same length.)

### Columns and rows

Because data frames are tabular data structures, we refer to the vectors they combine as *columns.*
The length of individual columns is equal to the number of *rows* of a data frame.
You can access this information with the `ncol()` and `nrow()` functions

### Creating new columns

We can also add new columns to an existing data frames by using the assignment operator:

```{r df-class, exercise.lines=20}
student_names <- c("Bill", "Jane", "Sarah", "Fred", "Paul")
math_scores <- c(80, 75, 91, 67, 56)
verbal_scores <- c(72, 90, 99, 60, 68)
students <- data.frame(student_names, math_scores, verbal_scores)

# add 'final_score' column
students$final_scores <- (students$math_scores + students$verbal_scores)/2
```

### Combining data frames 

Often we want to combine two or more data frames.

#### Column-wise combination

If we want to combine them side-by-side, we can "bind" them column-wise using the `cbind()` function:
```{r cbind}
a <- data.frame(letters = letters[1:2])
b <- data.frame(Letters = LETTERS[1:2])

cbind(a, b)
```

Note, however, that to allow for column-wise combination, the data frames need to have the same amount of rows!

```{r cbind-fail}
a <- data.frame(letters = letters[1:3])
b <- data.frame(Letters = LETTERS[1:2])

nrow(a)
nrow(b)

cbind(a, b)
```

#### Row-wise combination

If we want to stack two or more data frames on top of one another, we can "bind" them row-wise with the `rbind()` functions:

```{r rbind}
a <- data.frame(col = letters[1:2])
b <- data.frame(col = LETTERS[1:2])

rbind(a, b)
```

Note, however, that the `rbind()` function throws an error if the data frames you want to combine have different column names.

```{r rbind-fail}
a <- data.frame(letters = letters[1:2])
b <- data.frame(Letters = LETTERS[1:2])

rbind(a, b)
```


```{r rbind-ok}
a <- data.frame(letters = letters[1:2])
b <- data.frame(Letters = LETTERS[1:2])
names(b) <- "letters"

rbind(a, b)
```

#### Combining using a list of data frames

A very common scenario is when we have a list of data frames, and we want to bind them together.
We can do this by passing the list and our desired bind function as arguments to the `do.call()` function

```{r combine-do_call, exercise.lines=20}
results <- list()
# let's say here you're scraping 3 websites
results[[1]] <- data.frame(domain = "google",   url = "www.google.com")
results[[2]] <- data.frame(domain = "facebook", url = "www.facebook.com")
results[[3]] <- data.frame(domain = "twitter",  url = "www.twitter.com")

# and now we want to combine all 3 data frames
do.call(rbind, results)
```

Alternatively, we can use the `bind_rows` functions in the dplyr package.

```{r combine-bind_rows, exercise.setup = "combine-do_call"}
bind_rows(results)
```

## Control flow

Sometimes what you want to do depends on circumstances.
Implementing this behavior in code is called **control flow** because we control which of many potential streams of actions are executed.

Like in other programming languages, control flow is implemented with `if-else` statements.
Depending on whether a condition is true or false, we can execute different chunks of code.

### If

Let's start with `if`.
The logic is: "If something is true, do something."

```{r if-simple}
print_this <- FALSE
if (print_this) {
  print("Hello")
}
```

In the above code chunk, because `print_this` evaluates to FALSE, "Hello" is not printed.

Note that you can write everything inside the parentheses after `if` that evaluates to a single logical value (excluding `NA`).

```{r if}
if (0 == 1) {
  print("All numbers are equal!")
}
```


### If-Else

Sometimes you do one thing if a condition applies, and something else if it doesn't.

This can be implemented in R with `if`-`else` control flow:

```{r if-else}
if (0 == 1) {
  print("I escape the laws of math!")
} else {
  print("zero and one are not equal")
}
```

Note that there is the shorthand `ifelse()` function


```{r ifelse}
ifelse(0 == 1, "I escape the laws of math!", "zero and one are not equal")
```

### If, else, and else-if

If you think beyond binary options, you want to use `else if`:

```{else-if}
x <- 4
y <- 5
if (y > x) {
  print("`y` is greater than `x`")
} else if (x > y) {
  print("`x` is greater than `y`")
} else {
  print("`x` and `y` are equal")
}
```

## Loops

We use loops whenever we need to  execute the same chunk of code multiple times with some varying input. 

For example, we may use a loop whenever we have multiple Twitter accounts and we want to run sentiment analysis for tweets posted by each of them.

### `for`-loops

`for`-loops are probably the most common type of loop and are easily implemented in R

```{r for-loop}
vec <- 1:10
for (i in vec) {
	print(i)
}
```

### Anatomy of a `for`-loop

We can abstract from this to note that a for loop follows the following logic:

```
for (i in VECTOR) { 
  do something with i 
}
```

Now, let's have a look at the anatomy of a `for`-loop: 

- `for` tells R that you want repeat some code 
- you write the code you want to repeat *inside* the curly braces (i.e., between `{` and `}`)
- how often R evaluates this code and how is determined by the expression `(i in VECTOR)`
  - `VECTOR` is a vector (or list) with several elements
  - In each *iteration*, `i` is taking on the value of an element of `VECTOR`; first the value of the first element, then the value of the second element, and so on.
  
Consequently, when the code inside your loop (i.e., inside the curly braces) involves `i`,
in each iteration it is evaluated with the *current* value of `i`.
This allows to repeat code while varying `i`.

### `i` am a placeholder

Look at the following example and not that this code runs smoothly too.
This shows that instead of  `i`, you can use different word --- the word is is just a placeholder!

```{r fornumber}
for (number in 1:3) {
	print(number)
}
```

### Filling vectors/lists using for loops

There are many scenarios where you want to "collect" the results of individual for-loop iterations in a separate object.
Say we have a vector of integer numbers, we want to know if they are greater than 5, and we want to collect the result of this test in a new vector.


```{r for-loop-filling}
numbers <- 4:6
results <- logical(length = length(numbers))
for (i in seq_along(numbers)) {
  results[i] <- numbers[i] > 5
}
results
```

However, in the above example it is assumed that you know the length of the resulting object in advance.
Specifically, we know we will evaluate each element of `numbers` against the number five.
Because `numbers` has three elements, we know that we will generate 3 results.
Hence `results` must have three elements too.

### Appending values using for loops

If you don't know how long the vector you fill will be after all iterations have been completed, you can apply the following logic:

1. you create an empty vector/list
2. each iteration you generate a valid value, you "append" it by assigning it the the end of the vector/list 

Appending to the last position can be done by assigning a value to the position of the vector/list taht comes after the actual end of this vector, that is, at position `length(vector)+1`

```{r for-loop-appending}
numbers <- 4:6
results <- logical() # create zero-length/empty vector
for (i in seq_along(numbers)) {
  results[length(results)+1] <- numbers[i] > 5
}
results
```

You can verify yourself that because in this example 

- our `results` vector is empty in the begging (i.e., has length zero)
- and we iterate over $i = 1, 2, ...$

we can simply write:

```{r for-loop-appending-simple}
numbers <- 4:6
results <- logical()
for (i in seq_along(numbers)) {
  results[i] <- numbers[i] > 5
}
results
```

### Skipping and Stopping

Two important features of `for` loops is that you can skip iterations or stop them entirely if desired.
You skip an iteration using the keyword `next`

```{r for-loop-next}
for (i in 1:4) {
  # skip if `i` is less than 3
  if (i < 3) {
    next
  # otherwise print current value of `i`
  } else {
    print(i)
  }
}
```

You can interrupt a running fro loop with the `break` keyword:

```{r for-loop-break}
for (i in 1:4) {
  # stop if `i` is greater than 4
  if (i > 3) {
    next
  } 
  
  # otherwise print current value of `i`
  print(i)
}
```

## Functions

A function defines a procedure to return some output based on some inputs.

### Creating functions

We create functions in R using the `function()` function.

```{r function-no-arg}
hello <- function() {
  print("Hello")
}
```

### Calling functions

We can call this function as follows:

```{r  function-no-arg-call, exercise.setup = "function-no-arg"}
hello()
```

Note that because our `hello()` function does not have any arguments (i.e., parameters based on which it adapts its behavior), we do not pass it any values inside the parentheses. 

### Functions with arguments

This changes if we define our function in a way that it requires an input when called.
In our example, this input is called "name".


```{r function-one-arg}
hello <- function(name) {
  print(paste("Hello", name))
}
```

If we now call the function without passing an input, we get an error

```{r function-one-arg-fail, exercise.setup = "function-one-arg"}
hello()
```

So we have to pass a value to the `name` argument

```{r function-one-arg-success, exercise.setup = "function-one-arg"}
hello("Hauke")
```

### Returning results

In the functions above, we have just printed a value when calling them.
Usually, we use functions to compute something which the function is expected to return.

We can return values with the `return()` function that works only inside functions.

```{r function-return}
add <- function(x, y) {
  res <- x + y
  return(res)
}

add(1, 2)
```

*Note:* You could actually omit the return in the above function and instead write

```{r function-return-alt, exercise = FALSE}
add <- function(x, y) {
  x + y
}

add(1, 2)
```

But it aids the readability of your code *a lot* if you explicitly call `return()`.

### Returning early

`return()` can be called at any position in a function.
This is a particularly powerful feature when combined with control flow:

```{r function-return-anywhere}
divide <- function(numerator, denominator) {
  if (denominator == 0)
    return(NaN)

  if (numerator == 0)
    return(Inf)

  return(numerator/denominator)
}
```

