---
title: "Advanced web scraping techniques with the `rvest` R package"
author: "Theresa Gessler and Hauke Licht"
date: "last updated: `r format(Sys.Date(), '%e %B %Y')`"
description: >
  Learn about a number of advanced web scraping techniques with the `rvest` package:
  using sessions,
  interacting with HTML forms,
  handling login,
  and 
  managing user agents.
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r knitr, include=FALSE}
knitr::opts_chunk$set(
  # code chunk options
  echo = TRUE
  , eval = TRUE
  , warning = FALSE
  , message = FALSE
  , cached = FALSE 
  # learnr
  , exercise = TRUE
  , exercise.completion = TRUE
    # figs
  , fig.align = "center"
  , fig.height = 4
  , fig.width = 5.5
  , out.width = '50%'
)
```

```{r setup-invisible, include = FALSE, exercise = FALSE}
library(learnr)
```

## Introduction

In this tutorial, we show some important advanced web scraping techniques.

1. Using `sessions`
3. Interacting with HTML forms
2. Logging in
4. User agents

### R setup

We will use the following R packages:

```{r setup, exercise = FALSE}
library(rvest)
library(dplyr)
library(purrr)
```


## Using session

Thus far, we have relied on `xml2::read_html()` to read and parse websites' HTML code.
This is fine for many purposes.
But a more flexible way to interact with websites is through a browser session.
This has many advantages

- you can check the HTTP status code of your request
- you can pass configuration like "user agents"
- you can submit forms
- you can navigate, e.g., by following links

### Creating a session 

Sessions can be created in `rvest` with the `session()` function (`html_session()` before version 2.0.0).
The only required parameter is `url`, the address of the webpage to request.

**Try it yourself!**
First, have a look at https://scrapethissite.com/pages/simple/ which we will scrape in the example.
Next, create a session object using this URL.

```{r create-session}
url <- "https://scrapethissite.com/pages/simple/"
# To Do: create a session
sess <- ...
```

```{r create-session-invisble, exercise = FALSE, include = FALSE}
url <- "https://scrapethissite.com/pages/simple/"
sess <- session(url)
```

### Structure of a session object

If we inspect the return object assigned to `sess`, we see that it is a 'rvest_session' object.

```{r session-class, exercise.eval = TRUE, exercise.lines = 1}
class(sess)
```

Among others, objects of this class contain `httr` 'handle', 'request' and 'response' objects.

```{r session-str, exercise.eval = TRUE, exercise.lines = 1}
str(sess, 1)
```

The 'response' element is particularly useful because it

1. states the HTTP status code of the response, and
2. records the HTML code of the page your have requested (if the request had status '200')

```{r session-response, exercise.eval = TRUE, exercise.lines = 1}
sess$response
```

In addition, 'rvest_session' objects record the current URL

```{r session-url, exercise.eval = TRUE, exercise.lines = 1}
sess$url
```

### Parsing HTML code from a session object

Because the (parsed) HTML code of the requested page is associated with the 'rvest_session' object, we can call any `rvest` functions on it as we would do on the `xml_document` returned by `xml2::read_html()`.

For example:

```{r session-html_elements}
url <- "https://scrapethissite.com/pages/simple/"
sess <- session(url)

sess %>% 
  html_elements(".country-name") %>% 
  html_text(trim = TRUE) %>% 
  data.frame(country_name = .)
```

### Navigating

Session objects also allow navigating between pages.
To navigate to a new page, simply call `session_jump_to()` function on the session object and the target URL.

The session object then will keep track of the current URL and all previously visited URLs.
They can be reported by calling `session_history()`.

And because the session object 'remembers' the previous URL, we can go not only forward, but also backward.

```{r session-navigation}
url <- "https://scrapethissite.com/pages/simple/"
sess <- session(url)

# move forward
sess <- session_jump_to(sess, "https://www.scrapethissite.com/pages/forms")

# inspect history
sess$url
sess$back
session_history(sess)

# back to first page
sess <- session_back(sess)
sess$url

# and again back to second page
sess <- session_forward(sess)
sess$url
```


### Optional function parameters

Additional (optional) parameters can be passed to `session()`'s `...` argument, however, and they are forwarded to `httr::GET()`.
This makes configuring `rvest` sessions particularly easy.
Some of the most important features are 

- handling request headers (with `httr::add_headers()`,
- authentication (with `httr::authenticate()`,
- handling cookies (with `httr::set_cookies()`), and 
- and using proxies (with  `httr::use_proxy()`).

More on this later.

### *Advanced*: A peek under the hood

What does `rvest::session()`actually do?
The documentation (`?rvest::session`) is not particularly informative.
But we can learn something by looking at the function's *source code*:

It first creates a 'session' object calling the base R `structure()` function:
```r
session <- structure(
  list(
    # passed to argument 'handle' when calling httr::GET
    handle = httr::handle(url)
    # passed to argument 'config' when calling httr::GET
    , config = c(..., httr::config(autoreferer = 1L))
    # placeholder for URL that will be queried by rvest:::session_get
    , url = NULL
    , back = character()
    , forward = character()
    # placeholder for response objected that will be returned by httr::GET
    , response = NULL
    # ignore this
    , html = new.env(parent = emptyenv(), hash = FALSE)
  )
  , class = "session"
)
```

Next, it passes the 'session' object and the input URL to an internal function:

```r
rvest:::session_get(session, url)
```

This function makes a GET request using the `httr` package (see `?httr::GET`).
Specifically, 

1. it passes the URL to `httr::GET()`'s `url` argument
2. it passes the 'config' element of 'session' to `httr::GET()`'s `config` argument
3. it passes the 'handle' element of 'session' to `httr::GET()`'s `handle` argument
4. it makes the get request
5. it parses the response and adds information to the 'session' input object, which is returned enhanced with information like GET request response status

So `rvest::session()` is basically a wrapper around `httr::GET()` that eases interaction with request and response objects.

## Interacting with HTML forms

HTML forms allow collecting user input.

The page https://scrapethissite.com/pages/forms/ shows a simple example of a form --- in this case a *search bar*.
The search bar allows users entering key words to filter the data by that is shown in the table below the search bar.

Let's use this example to see how to interact with forms using `rvest`.

```{r form-session-invisible, exercise = FALSE, include = FALSE, eval = FALSE}
url <- "https://scrapethissite.com/pages/forms/"
sess <- session(url)
```

### Interacting with forms

The two most elemental ways of interacting with HTML forms are 

1. setting values
2. submitting the form

Before we can implement these in R, we first need to extract the form's corresponding R object and  understand its structure.

### Extracting forms

Forms included in a page can be extracted with the `html_form()` function.
Similar to `html_table()` (see tutorial "203-rvest-scraping-tables"), 
`html_form()` returns 

- a 'rvest_form' object if you input a single 'form' web element (i.e., a 'xml_node' object where `html_name(<obj>) == 'form'`), and 
- otherwise a list of 'rvest_form' objects.

**Try it yourself!** 
Extract all forms from https://scrapethissite.com/pages/forms/

```{r extract-forms}
url <- "https://scrapethissite.com/pages/forms/"
sess <- session(url)

# To Do: extract all forms
forms <- ...

# To Do: inspect the return object
str(forms, 1)
```

<!-- *Note:* if you apply `html_form()` directly to a 'form' web element (instead of a list of web elements as returned by `html_element()`), then this will return the 'rvest_form' object directly instead of in a list. -->

To interact with a specific form, we extract it from the list.
In this example, there is only one form that corresponds to the search bar displayed in the top of the page.

**Try it yourself!** 
Extract the frist form from https://scrapethissite.com/pages/forms/

```{r extract-a-form}
url <- "https://scrapethissite.com/pages/forms/"
sess <- session(url)

# get all forms
forms <- html_form(sess)

# To Do: extract the first form
a_form <- ...

# To Do: inspect the return object
str(a_form, 1)
```

```{r extract-a-form-invisible, exercise = FALSE, include = FALSE}
url <- "https://scrapethissite.com/pages/forms/"
sess <- session(url)
a_form <- html_form(sess)[[1]]
```

### Structure of a form objects

Let's have a look how form objects are structured!

```{r one-form, exercise.eval = TRUE, exercise.lines = 1}
str(a_form, 1)
```

If the form is named, its name can be accessed in the 'name' element.
The 'method' element notes the HTTP method.

Most important is the 'fields' element. It records the fields of the form.

```{r form-fields, exercise.eval = TRUE, exercise.lines = 1}
a_form$fields
```

Each element is a 'form_field' object with four elements:

```{r a-form-field, exercise.eval = TRUE, exercise.lines = 2}
a_field <- a_form$fields[[1]]
str(a_field, 1)
```

- 'type': the [form field type](https://www.w3schools.com/html/html_form_input_types.asp)
- 'name': the form parameter name (here, 'q' for "query")
- 'value': the value that is currently set on the field
- 'attr': attributes of the 'rvest_field' object

*Note:* 
Other field types may have additional elements.
It's always helpful to first inspect the fields of a form before interacting with it.

### Interacting with a form

Once we have extracted the form and understood its structure, we can interact with it.
The two most elemental interactions are 

1. setting values
2. submitting the form

#### Setting values

To set a value in a form, we use the `html_form_set()` function (previously `rvest::set_values()`).
For example, we can set a the query to a search term: "New York"
We pass this information as parameter--value pair to the function:

```{r fill-a-form, exercise.lines = 1}
a_form_filled <- html_form_set(a_form, q = "New York")
```

*Note:* If you want to set more than one form input parameter, simply add additional parameter--value pairs to the function call (e.g., `html_form_set(form, p1=v1, p2=v2, p3=v3)`).

```{r fill-a-form-invisible, exercise = FALSE, include = FALSE}
a_form_filled <- html_form_set(a_form, q = "New York")
```

We can look at the value of the 'q' text field to verify that our query has been set:

```{r fill-a-form-check-value, exercise.lines = 1}
a_form_filled$fields$q$value
```

#### Submitting a form

To submit a form, we require the current session object and the filled form.
We then pass these objects to the `session_submit()` function:

```{r submit-a-form, exercise.lines = 1}
sess <- session_submit(sess, a_form_filled)
```

*Note:* Because by submitting the form the page content may change, we "update" the session object by simply overwriting it .

```{r submit-a-form-invisble, exercise = FALSE, include = FALSE}
sess <- session_submit(sess, a_form_filled)
```

We can verify that the submission has worked by extracting the results shown in the searchable data table:

```{r submit-a-form-check}
sess %>% 
  html_element("table") %>% 
  html_table() %>% 
  pull(1) %>% 
  table()
```

Great! The names of all listed Hockey team contain the term "New York".

## Login

Simple sign-on login is easy to handle once you know how to handle HTTP forms with `rvest`.
We simply

1. locate the login form,
2. pass our user information (email/user name and password), and
3. submit the filled login form

You can try this on https://www.stealmylogin.com/demo.html

```{r login-example}
# create session
url <- "https://www.stealmylogin.com/demo.html"
sess <- session(url)

# 1. locate the login form
login_form <- html_form(sess)[[1]]

# 2. pass user info

# what fields and types ?
purrr::map_chr(login_form$fields, "type")

# set values
login_form <- html_form_set(
  login_form, 
  # pass values to form fields
  "username" = "test.user@gmail.com", 
  "password" = "123456"
)

# check
purrr::map_chr(login_form$fields, "value")

# 3. submit 
sess <- session_submit(sess, login_form)


# Inspect result:

# was login successful?
httr::status_code(sess) # should be 200

# what info was posted?
URLdecode(rawToChar(sess$response$request$options$postfields))

# where have we been redirected?
sess$url
```


## User agents

When making HTTP requests, we send information about who is making the the request.
That is, we tell the server what **user agent** we are.

Because to scrape content from websites, `rvest` relies on `httr` for making HTTP requests, which, in turn, makes these request using the [`curl` program](https://curl.se/), by default we send this as user agent information when we execute `httr::GET` or `rvest::session()`.

```{r default-user-agent, exercise.lines = 1}
sess$response$request$options$useragent
```

The same applies when we read and parse the HTML code of a website with `xml2::read_html()` because, to read data from a connection, under the hood  `xml2::read_html()` uses `curl` as well.

There are many reasons why we want to overwrite this default user agent information:

- transmit information about who you are and why you automate acessing their server
- avoid getting your user agent banned when scraping

### Specifying the user agent

Remember how we said that `rvest::session()` simply wraps `httr::GET()` and this makes configuring interactions with websites really easy.
Handling user agents is a perfect point in case.
When creating a session object, we simply pass a user agent object created with `httr::user_agent()` to the `...` argument of `rvest::session()`:

```{r setting-users-agent}
ua <- "Mozilla/5.0 (Macintosh; Intel Mac OS X 10_15_7) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/91.0.4472.114 Safari/537.36"
sess <- session(url, httr::user_agent(ua))

# check that this was successfull
is.null(sess$config$options$useragent) # should be FALSE
sess$config$options$useragent == ua # should be TRUE
```


