---
title: "Downloading files"
author: "Theresa Gessler"
output: learnr::tutorial
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(rvest)
knitr::opts_chunk$set(echo = FALSE)
```

### Introduction

*Last revised: `r Sys.Date()`*

Similar to scraping tables, downloading files is pretty easy. Actually, it is one of the tasks for which we do not even need `rvest`! Essentially, downloading files works with base-R and all that our knowledge of scraping does is make it easier for us to collect the links and to think through how to download files. Still, I have preloaded `rvest()` to help you with collecting file paths.

### Downloading Files in R

The central command for us is `download.file()`. If you have never used it, check [its documentation](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/download.file.html){target="_blank"}.

Essentially, you need to specify:

-  the **url** you are downloading from,
- the **destfile** (destination file: where and under which name your file will be saved)
- and the **method** of downloading

The method of downloading works a bit different on each operating system, so I would again recommend that you check the [documentation](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/download.file.html){target="_blank"} if the default does not work for you. The same holds for the **mode** of downloading.

To practice, we download the [APSA Diversity and Inclusion Report](https://www.apsanet.org/Portals/54/diversity%20and%20inclusion%20prgms/DIV%20reports/Diversity%20Report%20Executive%20-%20Final%20Draft%20-%20Web%20version.pdf){target="_blank"}. 


Let me just describe what the following code chunk does: 

- It specifies the path to the file we want to download
- It generates a filename 
    - the command `basename()` keeps only the last part of the path of the pdf, namely the filename
    - you can also use the destfile specification to specify a full file path
 - it sets the mode of downloading - for me the correct mode is "wb" but something different might work for you.

**Since this tutorial does not run via your normal R console you won't see the downloaded files so you might choose to execute this specific exercise directly in R instead!**  If you try it out in your R console, the file should appear in your working directory.

```{r practice,exercise=TRUE}
download.file("https://www.apsanet.org/Portals/54/diversity%20and%20inclusion%20prgms/DIV%20reports/Diversity%20Report%20Executive%20-%20Final%20Draft%20-%20Web%20version.pdf",
    destfile=basename("https://www.apsanet.org/Portals/54/diversity%20and%20inclusion%20prgms/DIV%20reports/Diversity%20Report%20Executive%20-%20Final%20Draft%20-%20Web%20version.pdf"),
    mode="wb")
```





### Downloading Files within a webscraping routine

To be honest, situations in which you have to download lots of files rather than HTML pages are scarce, still, they happen. Let's pretend we want to download the [Congressional Record of the ongoing session](https://www.congress.gov/congressional-record/116th-congress/browse-by-date){target="_blank"} in its [beautiful original layout as a pdf](https://www.congress.gov/116/crec/2019/06/28/CREC-2019-06-28.pdf){target="_blank"}. There are several steps to this:

- find the appropriate selectors and scrape all links to pdf from the [page of the Congressional Record of the ongoing session](https://www.congress.gov/congressional-record/116th-congress/browse-by-date){target="_blank"} - this is really a scraping task like before
- create destination filenames, using the command `basename()` - just as we did before
- use `mapply()` to apply the `download.file()` function to both the file name and the destination file; 
    - if you use `sapply()` or a similar function, only the url will vary but the document will always be saved under the same name - `mapply()` can go over multiple vectors (in our case: URLs and filenames)
    - it works like this: `mapply(function, vector1, vector2, otherarguments)` - note that the parameters are in a different order than for `sapply(vector, function,otherarguments)`
    - if it is easier for you, you can also write a function that wraps `download.file()` with the correct mode

**Pro-Tip: Don't download all files since we won't use them - just cut the vector of paths before you proceed. Again, since this tutorial does not run via your normal R console you won't see the downloaded files so you might choose to execute it directly in R instead!**

```{r congress, exercise=TRUE,exercise.lines=10}

```

```{r congress-solution}
paths<-read_html("https://www.congress.gov/congressional-record/116th-congress/browse-by-date") %>% html_nodes("td:nth-child(6) a") %>% html_attr("href")
paths<-paths[1:5]
filenames<-basename(paths)
paths<-paste0("https://www.congress.gov",paths)
mapply(download.file,paths,filenames,mode="wb")

# OR with a wrapper function:
download_files<-function(paths,filenames){
	download.file(url=paths,destfile=filenames,mode="wb")
}
mapply(download_files,paths,filenames)
```

