---
title: "Scraping Tables"
output: learnr::tutorial
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

Compared to text, tables are significantly easier to scrape: Using html_table() we can just extract all tables contained in the page and then select those that are relevant for us. Or, we can again use html_nodes to select specific tables.

### Scraping all tables

The single most important specification is the *fill* parameter. If you specify fill as true, rvest will automatically fill rows with fewer than the maximum number of columns with NAs. This is useful because it allows you to scrape messy tables with inconsistent numbers of cells per row.

Try it out on [wikipedia's list of the tallest buildings](https://en.wikipedia.org/wiki/List_of_tallest_buildings.

```{r fill, exercise=TRUE}
"https://en.wikipedia.org/wiki/List_of_tallest_buildings"
```

```{r fill-hint}
read_html("https://en.wikipedia.org/wiki/List_of_tallest_buildings") %>% html_table()
read_html("https://en.wikipedia.org/wiki/List_of_tallest_buildings") %>% html_table(fill=T)
```

If you assign the result to an object, the result will be a list.
You can extract specific tables from this list by subsetting the list with two brackets. Or, if you want to proceed in a piping-chain, you can use the command extract2(), adding the number of table in brackets (the command name is no typo - extract without the 2 works for vectors). Try both variants for extracting the second table from the page we downloaded.

```{r extract, exercise=TRUE}
"https://en.wikipedia.org/wiki/List_of_tallest_buildings"
```

```{r extract-hint}
tables<-read_html("https://en.wikipedia.org/wiki/List_of_tallest_buildings") %>% html_table(fill=T)
tables[[2]]
tables %>% extract2(2)
```

### Selecting specific tables

Alternatively, you can select specific tables upfront. This is for example useful when you scrape different pages that all contain the relevant information but potentially in a different order - think wikipedia articles about artists or writers that all contain a list of their work but may also contain other optional tables beforehand.

In this case, you can just use html_nodes() to extract the relevant table. 