---
title: "Interactive learning with `learnr` tutorials"
# author: "Theresa Gessler and Hauke Licht"
# date: "last updated: `r format(Sys.Date(), '%e %B %Y')`"
description: >
  Learn how to work with itnteractive `learnr` tutorials
   and
   about the specifics of tutorials included in the `learn2scrape` package.
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r knitr, include=FALSE}
knitr::opts_chunk$set(
  # code chunk options
  echo = FALSE
  , eval = TRUE
  , warning = FALSE
  , message = FALSE
  , cached = FALSE
  , exercise = TRUE
)
```

```{r setup, include=FALSE}
library(learnr)
```

<!-- Welcome to web scraping with learn2scrape! -->
<!-- This is a first attempt at making this class more interactive and fun. -->
<!-- If you have any feedback, let me know in the next days [or after the course via e-mail](mailto:gessler@ipz.uzh.ch)! -->
<!-- Also, if you are interested in scraping, feel free to [watch the repository on Github](https://github.com/theresagessler/learn2scrape) to get updates on new elements. -->

The tutorial has explanations but there are many interactive parts where you will be asked to write code. 
Some notes about the format:

- If you don't know the answer, you can click on *Hint* (where available) or the *Solution* Button to see the solution
- You can test your code by clicking on *Run Code*
- Since there is no environment storing R object (like in RStudio), you have to 
  1. directly call the objects to inspect them, 
  2. call `str(<obejct>)` to inspect their structure, or 
  3. use `ls()` to see what is in your environment when you execute the chunk
- Be cautious about the *Start over* Button: Clicking on it will erase all the code you wrote thus far.

Unlike in a normal R markdown document, all chunks are independent. 
That means: If you create something in one chunk, it won't be available in the other chunks. 
Therefore, you will sometimes have to repeat code from previous exercises but I have tried to keep this to a minimum. 
Either use it as an opportunity to repeat or copy-paste - learnr saves your answers.

##### *Note*

Some of the exercises will use [piping](https://magrittr.tidyverse.org/) (i.e., the `%>%` operator) since it just makes scraping so much easier. 
If you are not familiar with this or any other code, just let me know.

<!--If you prefer using [swirl](https://swirlstats.com/){target="_blank"}, this first lesson is also available as a swirl-course. As things get more complex later on, I'll ask you to switch to the browser-version! -->
