---
title: "Scraping data from APIs"
author: "Theresa Gessler and Hauke Licht"
date: "last updated: `r format(Sys.Date(), '%e %B %Y')`"
description: >
  Learn the basics of querying data from *Application Programming Interfaces* (APIs)
  with the `httr` R package.
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---


```{r setup-invisible, include = FALSE}
library(learnr)
```


```{r knitr, include=FALSE}
knitr::opts_chunk$set(
  # code chunk options
  echo = TRUE
  , eval = TRUE
  , warning = FALSE
  , message = FALSE
  , cached = FALSE 
  , exercise = TRUE
  , exercise.completion = TRUE
)
```

## Introduction

In this tutorial, you will learn the basics of querying APIs using R.
We will rely on the *Dad Jokes* API as a running example: https://icanhazdadjoke.com/api

We use the following packages:
```{r, exercise = FALSE}
library(httr)
library(dplyr)
```
```{r setup, exercise = FALSE,include=F}
library(httr)
library(dplyr)
endpoint <- "https://icanhazdadjoke.com"
searchendpoint <- "https://icanhazdadjoke.com/search"

```

## Read the documentation 

Go to https://icanhazdadjoke.com/api and familiarize yourself with the API.
Then answer the following questions.

```{r endpoint-quiz, exercise = FALSE, echo = FALSE}
quiz(
  caption = "API endpoints",
  question(
    "What is the correct endpoint to randomly query one joke from the *Dad Jokes* database?",
    answer("https://icanhazdadjoke.com/random", message = "Almost!"), 
    answer("https://icanhazdadjoke.com/api", message = "No, this is the URL of the API documentation."),
    answer("https://icanhazdadjoke.com", correct = TRUE),
    allow_retry = TRUE
  )
  , 
  question(
    "What is the correct endpoint to search for dad jokes?",
    answer("https://icanhazdadjoke.com/search", correct = TRUE), 
    answer("https://icanhazdadjoke.com/j", message = "No, this is the endpoint to query jokes by ID."),
    answer("https://icanhazdadjoke.com/find", message = "Almost!"),
    allow_retry = TRUE
  )
)
```

## Simple example

Lets use the endpoint that returns a random joke to learn how to query an API in R.
The correct endpoint is https://icanhazdadjoke.com
We first assign the endpoint URL to an object called 'endpoint'.
  
```{r random-endpoint, exercise = FALSE}
endpoint <- "https://icanhazdadjoke.com"
```


### Making a simple request

No we can make a GET-request to this API endpoint using the `GET()` function from the `httr` package.
If you check the documentation of this function (type `?httr::GET`), you'll see that it expects the endpoint URL as input to the frist argument.
Since we just want to query a random joke, and this endpoint accepts no parameters (more on this later), we simply call `GET()` with the endpoint URL as single input:

```{r random-joke-html, exercise.setup = "random-endpoint"}
resp <- GET(endpoint)

# inspect
resp 
```


```{r random-joke-html-invisible, exercise = FALSE, echo = FALSE, exercise.setup = "random-endpoint"}
resp <- GET(endpoint)
```


If you print the return object `resp`, you see some important information:

- The [*response status*](https://developer.mozilla.org/en-US/docs/Web/HTTP/Status) is 200. This is the numerical code signalling a "success". Great!
- The [*content type*](https://developer.mozilla.org/en-US/docs/Web/HTTP/Basics_of_HTTP/MIME_types/Common_types) of the returned data is text/html. So the data is formatted as HTML. This is important when we want to **parse** the response to make it usable in R.

### Parsing

```{r random-joke-html-parse, exercise.setup = "random-joke-html-invisible", exercise.lines = 3}
# parse returned data
content(resp, "parsed") # this is an xml2/rvest 'html_document' object
```

The default return format of the *Dad Jokes* API is HTML for this API.
Most other APIs, however, return data in JSON format.

We can instruct the API to return the data in another format by by adding a header to our query.
The 'Accept' header tells the API that it should return data in JSON format.

```{r random-joke-json, exercise.setup = "random-joke-html"}
resp <- GET(endpoint, config = add_headers(Accept = "application/json"))

# if we directly parse the JSON, we get an R 'list' object
content(resp, "parsed")
```

Note that we could also simply specify `config = accept_json()`:

```{r, exercise = FALSE, eval = FALSE}
resp <- GET(endpoint, config = accept_json())
```

<br/>

**Try it yourself!**
Check the [docu](https://icanhazdadjoke.com/api#api-response-format) and try to obtain the response in plain text format

```{r text, exercise.lines = 3,eval=F}
resp <- GET(endpoint, config = ...)
```

## API queries with parameters

The above exampel was quite simple because the endpoint we've queried did not accept any parameters.
But most APIs have parameters that allow clients to control what data is returned from an API.

In this sense, APIs function just like R functions: you pass a value to a pre-defined argument to manipulate the behavior of the API and to control what data is returned.

This is also why its so important to familiarize yourself with the documentaion when you work with an API: it states what query parameters there are and what kind of inputs they accept/expect.

### Example: Passing a search query

The *Dad Jokes* API also provides a **search endpoint**.
This is quite common for APIs that provide text data.
The search endpoint is https://icanhazdadjoke.com/search

```{r search-endpoint, exercise = FALSE}
searchendpoint <- "https://icanhazdadjoke.com/search"
```

It accepts three parameters:

- `term`: search term to use (default: list all jokes)
- `page`: which page of results to fetch (default: 1)
- `limit`: number of results to return per page (default: 20) (max: 30)

### Search example

We will ignore the last two parameters for the moment and just use the `term` parameter to search for jokes that include the term 'drink'.
With `httr::GET()`, API parameters need to be passed to the function parameter `query` as a list.
So we need to add `query = list(term = 'drink')` when calling get

```{r joke-search-drink, exercise.setup = "random-joke-html"}


```

<details>
<summary>**_Example Solution (click)_**</summary>
```{r , exercise = FALSE}
resp <- GET(
  searchendpoint, 
  query = list(term = 'drink'), 
  config = accept_json()
)
```
</details>

```{r joke-search-drink-invisible, exercise = FALSE, include = FALSE, exercise.setup = "random-joke-html"}
resp <- GET(
  searchendpoint,
  query = list(term = 'drink'),
  config = accept_json()
)
dat <- content(resp, "parsed")
```

We can verify that the key-value pair term and 'drink' has been added to the API request by checking the URL:

```{r joke-search-drink-url, exercise.setup = "joke-search-drink-invisible", exercise.lines = 1}
resp$request$url
```

We can now inspect the parsed response:

```{r joke-search-drink-structure, exercise.setup = "joke-search-drink-invisible", exercise.lines = 1}

```
<details>
<summary>**_Example Solution (click)_**</summary>
```{r , exercise = FALSE}
dat <- content(resp, "parsed")

str(dat, 1)
```
</details>


Inspecting its structure, we see that the API has returned a total 3 jokes (see list element 'total_jokes') and that there is an element called 'results.'
This element records the jokes matching our key term:

```{r joke-search-drink-results, exercise.setup = "joke-search-drink-invisible", exercise.lines = 1}
dat$results
```

## Querying multiple jokes

To query multiple jokes, we will again use the *search endpoint*: https://icanhazdadjoke.com/search

Instead of searching for jokes that include a certain key word, we will now try to download *all* jokes in the the *Dad Jokes* database.
This example helps to clarify some best practices:

- how to implement "pagination"
- how to use for-loops to avoid repetitive code
- how to use custom R functions to make your code robust

### Querying any joke

The first thing to note is that we can pass an empty query string '''' to the `term` parameter.
Because all joke texts will match this query term, the API will return *all* jokes.

*Note:* Since this is the default search term, we can simply omit it from our API request.

In this way we can iterate over pages to extract as many dad jokes as we'd like.

### Initial query

To see how this can be done, let's first make an initial query to the search endpoint and parse the returned JSON data into an R object:

```{r joke-search-all-parsed, exercise.setup = "random-joke-html"}
resp <- GET(searchendpoint, config = accept_json())

dat <- content(resp, "parse")
str(dat, 1)
```

```{r joke-search-all-parsed-invisible, exercise = FALSE, echo = FALSE, exercise.setup = "random-joke-html"}
resp <- GET(searchendpoint, config = accept_json())
dat <- content(resp, "parse")
```

As you see, in addition to 20 jokes (recorded in the 'results' list element), the returned data indicates 

1. from which page we have currently obtained jokes (page 1, by default),
2. what is the next page number, and 
3. how many jokes there are in total

That's great because we can use this information to extract more than the first 20 jokes!

### Plan of attack

Let's first brainstorm how we'd do this:

1. We need to find some way to obtain jokes' text (and maybe also ID) from a single query's response.
2. Next, we need to find a way to iterate over all pages until we have collected all jokes

### Step 1: Extracting jokes data

Let's start with the first point!
We will use the data returned by our first query.
As seen above, jokes are in the list element called 'results'.

```{r search-jokes-all-response-structure, exercise.setup = "joke-search-all-parsed-invisible"}
# 'results' is a list of lists
str(dat$results, 1)
# all sub-lists have two elements
lengths(dat$results)
```

Let's check how a sub-list looks like

```{r search-jokes-all-response-sublist, exercise.setup = "joke-search-all-parsed-invisible", exercise.lines = 1}
dat$results[[1]]
```

We already know this from the random joke endpoint:
The returned data records jokes' texts and their IDs.

We can obtain this data from the 'results' list element using the `dplyr` function `bind_rows()`. 
This results in a data frame with two columns: 'id' and 'joke'

```{r search-jokes-extact-jokes, exercise.setup = "joke-search-all-parsed-invisible", exercise.lines = 2}
jokes <- bind_rows(dat$results)
head(jokes, n = 3) # show first 3 rows
```

In base R, we could instead write:

```{r , exercise = FALSE, eval = FALSE}
jokes <- lapply(dat$results, as.data.frame.list, stringAsFactor = FALSE)
jokes <- do.call(rbind, jokes)
```

### Step 2: Pagination

Next, we want to determine the next page to query data from.
As seen above the parsed API response records this information in a list element called 'next_page.'

```{r search-jokes-next-page-value, exercise.setup = "joke-search-all-parsed-invisible", exercise.lines = 1}
dat$next_page
```

So we want to use this information to query jokes from the next page.
How do we do this?
We add query *parameters* to our API request!

**Remember:** the three valid parameters of the search endpoint are 'term', 'limit' and 'page'.
So if after a current request, we get the value from the 'next_page' field of the response data and pass this value to the 'page' parameter in the following query, we can iterate over all available pages.

So the R (pseudo) code would be:
```{r , exercise = FALSE, eval = FALSE}
# STEP 1
first_request <- GET(searchendpoint)
params <- list(page = first_request$next_page)

# STEP 2
second_request <- GET(searchendpoint, params)
params <- list(page = second_request$next_page)

# STEP K 
...
```

#### Using a `for`-loop

We can solve this with a `for`-loop!
However, a for loop requires us to pre-define the page number!

Fortunately, we can compute this from the information returned by the first request:
Remember that in the returned data there was a field called 'total_pages' --- this is exactly what we need!

```{r search-jokes-n-pages}
# extract total number of pages and construct integer sequence
pages <- seq_len(dat$total_pages)

# inspect
length(pages)
range(pages)
```

#### The for loop anatomy

Now we can think about the code of the for loop:

We want to iterate over pages. 
So we write `for (i in pages) { ... }`

We replace `...` with the code we want to execute: 

```{r exercise = FALSE, eval = FALSE}
# make request for page `i`
resp <- GET(searchendpoint, query = list("page" = i), accept_json())

# parse response
dat <- content(resp, "parsed")

# extract jokes data (bind row-wise)
dat <- bind_rows(dat$results)
```

Note that in each iteration, we want to collect the result assigned to `dat`.
We can do this by appending to a list we create beforehand:

```{r exercise = FALSE, eval = FALSE}
# define list collecting jokes
all_jokes <- list()

# iterate over pages
for (page in pages) {
  
  ... 
  # append current data jokes data
  all_jokes[[page]] <- dat
}
```

### Combining results 

After completing all iterations, `all_jokes` will be a list of data frames.
You probably know (e.g. from the R basics tutorials) that we can again use `bind_rows()` to bind these data frames (by row) to one long data frame

### Being patient

We need to add one more essential ingredient to our code: **Patience**!
If we run the `for`-loop just like that, it will send a lost of request within just a few second.
API providers don't like that!

So we need to pause between subsequent request.
Pausing can be imlemented using the system utility function `Sys.sleep()`.
This function accepts a number (double or integer) as single input which determines how long the system pauses (in seconds) before returning.

The following code illustrates this

```{r pause}
# get time at start
st <- Sys.time()
# pause for 3.33 seconds
Sys.sleep(3.33)
# print difference to start time when done
Sys.time()-st
```

*Hint:* We can randomize how long we pause by sampling a single number from a probability distribution such as the uniform distribution (implemented as `runif()` in R).

### Putting everything together

Now try to put everything together.
It will take quite some time to complete all iterations. 
So we recommend you copy-paste it to an R file in your editor and run it independently.

```{r search-jokes-pagination-solution, exercise = FALSE, eval = FALSE}
# make initial query



# extract total number of pages and construct integer sequence



# define list collecting jokes



# iterate over pages
for (page in pages) {


  
  # query API


  # parse response


  
  # extract jokes data


  
  # pause for 1-2 seconds


}

# Finally: combine jokes data in data.frame

```

<details>
<summary>**_Example Solution (click)_**</summary>
```{r , exercise = FALSE,eval=F}
# make initial query
resp <- GET(endpoint, accept_json())
dat <- content(resp, "parsed")

# extract total number of pages and construct integer sequence
pages <- seq_len(dat$total_pages)
# inspect
length(pages); range(pages)

# define list collecting jokes
all_jokes <- list()

# iterate over pages
for (page in pages) {
  # print current page number
  message("\b\rpage ", page)
  
  # query API
  resp <- GET(endpoint, query = list("page" = page), accept_json())
  # parse response
  dat <- content(resp, "parsed")
  
  # extract jokes data
  jokes[[page]] <- bind_rows(dat$results)
  
  # pause for 1-2 seconds
  Sys.sleep(runif(1, 1, 2))
}

# Finally: combine jokes data in data.frame
all_jokes <- bind_rows(all_jokes)
```

</details>


### Making things more efficient

The above code can be made more efficient in two ways:

1. instead of starting to iterate from the first page, we could actually start from the second page. Why? because to determine the total number of pages, we have queried data from the first page.
2. We could increase the number of jokes per page to 30 (the API-internal maximum number). In this way, we would need to iterate over fewer pages, hence making fewer API request. This would reduce the amount of web traffic we generate as well as the total amount of seconds we pause between subsequent API requests.















